### 2021-05-17 manuscript/v8_submission_finalization/assignment_files/manuscript_tex4submission.tex
final adjustments for manuscript

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{Causal graph analysis of COVID-19 observational data in German districts reveals effects of determining factors on reported case numbers} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Edgar Steiger\textsuperscript{1*},
Tobias Mußgnug\textsuperscript{1},
Lars Eric Kroll\textsuperscript{1}%,
%Name4 Surname\textsuperscript{2},
%Name5 Surname\textsuperscript{2\ddag},
%Name6 Surname\textsuperscript{2\ddag},
%Name7 Surname\textsuperscript{1,2,3*},
%with the Lorem Ipsum Consortium\textsuperscript{\textpilcrow}
\\
\bigskip
\textbf{1} Central Research Institute of Ambulatory Health Care in Germany (Zi),
Salzufer 8, D-10587 Berlin, Germany
\\
%\textbf{2} Affiliation Dept/Program/Center, Institution Name, City, State, Country
%\\
%\textbf{3} Affiliation Dept/Program/Center, Institution Name, City, State, Country
%\\
\bigskip

@@

We used a directed acyclic graph (DAG)
\cite{schipf_directed_2011, textor_robust_2017} as a tool to analyze
the causal relationships between several exposures and SARS-CoV-2
spread. To get an overview on published associations, a scoping review
was conducted from 20th to 22nd of May 2020 within Pubmed and Google
scholar. Restrictions were applied to English and German language and
the publication date in the last one year. The following search terms
were applied to abstracts and title in Pubmed (``COVID-19'' OR
``COVID19'' OR ``Corona'' OR ``Coronavirus'' OR ``SARS-CoV-2'') and
connected separately in each case with the exposure variables
(``mobility'', ``public awareness'', ``awareness'', ``google
trends'',``ambient temperature'', ``temperature''). For ``mobility'', we
analyzed \(n=8\) studies, \(N=103\) were scanned in Pubmed, together
with the first ten pages (100 results) in Google scholar
(``awareness''/``public awareness''/``google trends'' \(n=9\),
\(N=215\); ``temperature''/``ambient temperature'' \(n=16\), \(N=235\)).
We integrated these findings where possible into the construction of our
DAG, which can be seen in Fig~\ref{fig:dag-covid-19}.

@@

\begin{figure}[!h]
\caption{{\bf DAG of determinants of reported COVID-19 cases on the district level.}
Unobserved variables are light gray, variables marked with an asterisk (*) are confounded by weekday/holiday.}
\label{fig:dag-covid-19}
\end{figure}

@@

We collected and aggregated data on reported COVID-19 cases, regional
socio-demographic factors, weather, and general mobility on district and
state level in Germany for the period of 15 February 2020 to 8 July
2020. Our observation period for the outcome consisted of all dates from
20 February 2020 to 8 July 2020 (\(T=140\)), since we used a lag of 5
days for all confounders. We did not exclude any states or districts
(\(K=401\)). We analyzed the daily reported number of new cases as
outcome (\(K\cdot T=56{\ }140\) observations). The set of possible
predictors was derived from our causal DAG (see Table~\ref{tab:model-vars} and Fig~\ref{fig:dag-covid-19}). Due to
modelling and data limitations, some of the predictors were unobserved
or were modelled as a construct consisting of several variables. For our causal graph analysis, we computed adjustment sets separately for all observed exposures within the DAG (if the respective exposure was identifiable within the DAG causal analysis framework).

@@

We analyzed the DAG from Fig~\ref{fig:dag-covid-19} with the R Software
\cite{rsoftware} and the R packages \texttt{dagitty} (formal
representation of the graph and minimal adjustment sets
\cite{textor_robust_2017}) and \texttt{pcalg} (for finding an optimal
adjustment set \cite{pcalg}). For the defined exposures and the outcome
``Reported new cases of COVID-19'', we computed the minimal and optimal
adjustment sets. Since it was possible that these sets contained unobserved variables that needed to be left out of the regression model, we chose the valid set with the lowest AIC (see next section) to estimate the final total causal effect from exposure to outcome.

@@

In the observational period, the number of daily reported COVID-19 cases increased till the end of March/beginning of April and continually decreased afterwards till the beginning of June 2020 with a slight increase and decrease afterwards (Fig~\ref{fig:cases-log}A). On the other hand, the (log-)ratio of reported cases over active cases decreased steeply till the mid of April and increased steadily afterwards with a slight decrease close to the end of the observation period (Fig~\ref{fig:cases-log}B). Both figures examplify a considerable variation among the districts (light blue points are individual district's data).

\begin{figure}[!h]
\caption{{\bf Temporal and district level variation of outcome (log-scale).}}
\label{fig:cases-log}
\end{figure}

In Germany, we observed a rebound in mobility after the initial political measures, reductions in incident cases were associated with a diminishing public interest in COVID-19, and temperatures were overall increasing (cf.~Fig~\ref{fig:timeseries}); with correlations between temporal progression and mobility in retail and recreation \(r_{A,B}=\) 0.02, awareness (``Searches corona'') \(r_{A,C}=\) -0.3, and temperature \(r_{A,D}=\) 0.8.

\begin{figure}[!h]
\caption{{\bf Temporal variation of outcome and main
determinants.}}
\label{fig:timeseries}
\end{figure}

@@

We list the results of our causal analysis for the effects of different exposure variables in Table~\ref{tab:causaleffects}. The estimates are multiplicative rates of increase/decrease for a one unit increase of the respective variable: Values above 1 lead to an increase, below 1 to a decrease of the infection rate. To put these estimates into perspective, Fig~\ref{fig:effectsrange} shows the relative causal effect of the different exposure variables on the number of reported COVID-19 cases on a range of sensible values of the exposure variables (95 percent quantiles of data points).

@@

\begin{figure}[!h]
\caption{{\bf Relative causal effects of exposures.}}
\label{fig:effectsrange}
\end{figure}

Within our framework, we saw very different effects for individual mobility variables. For mobility in retail/recreation, an increase of 1 percent point mobility compared to the reference period (03 January to 06 February 2020) leads to an increase of the daily reported case number by about 0.11 percent. Similarly, mobility on workplaces showed an effect of 0.33 increase in case numbers for every 1 percent point increase in mobility, while mobility on transit stations showed an effect of 0.26 increase in case numbers for every 1 percent point increase. Contrarily, the remaining three mobility variables showed negative effects on the number of reported COVID-19 cases. An increase of 1 percent point mobility for the areas of grocery/pharmacy leads to a decrease in the reported case number by approximately 0.23 percent, while increased mobility of 1 percent point within parks leads to a decrease in the reported case number by approximately 0.03 percent, and finally an increase of 1 percent point in residential mobility leads to a decrease by approximately 0.97 percent. Fig~\ref{fig:effectsrange} shows the effects of mobility on a range of possible values. Thus, we expect an increase of daily cases by approximately 7.8 percent if mobility in workplaces reaches baseline levels of 0 percent difference to the reference period. On the other hand, an increase of mobility for residential areas by 10 percent points compared to the reference period leads to a reduction of the infection rate by approximately 1.8 percent.

``Awareness'' had two opposite effects on the outcome in our DAG. Awareness measured by Google searches for \emph{corona} had a positive effect on the number of reported cases. An one percent point increase of the state's Google searches (relative to other states and the observation period) leads to an increase of approximately 0.89 percent. For example, if a district shows 10 percent points more relative searches for \emph{corona} than another one, we expect approximately 9.3 percent more infections for this district after 5 days. \emph{COVID-19 burden} (reported number of cases on day of exposure) affected the outcome negatively, where every additional daily case in the district leads to a 0.2 percent decrease in newly reported case numbers. The corresponding plot in Fig~\ref{fig:effectsrange} visualizes this relationship: For a local outbreak with 20 daily cases as COVID-19 burden, we estimate as total causal effect a subsequent reduction of infection rate by 3.9 percent.

Within our model, we observed effects of temperature and all other weather variables. Every increase of 1 degree Celsius in temperature leads to a reduction of the daily reported case numbers by approximately 0.95 percent. On the other hand, we found an increasing effect of rainfall: One millimeter (=1 liter per square meter) more rainfall leads to an increase of reported case numbers by approximately 1.21 percent. We observe effects for humidity and wind as well (higher humidity and stronger wind leading to more cases). In perspective (Fig~\ref{fig:effectsrange}), with temperature we expect an increase by approximately 21 percent at a daily average temperature of \(0^{\circ}C\) compared to a day with \(20^{\circ}C\). For rainfall, we expect on a rainy day with 10 mm rainfall a corresponding increase of the infection rate by approximately 12.8 percent compared to a day with no precipitation.

@@

The code and resources for our analysis are available on Github, we invite other researchers to replicate our analysis with different assumptions using the files provided in the repository\footnote{The repository is located here: https://github.com/zidatalab/causalcovid19} of the article.

@@

\bibliography{mybibfile}

>>>>

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{Causal graph analysis of COVID-19 observational data in German districts reveals effects of determining factors on reported case numbers} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Edgar Steiger\textsuperscript{1*},
Tobias Mussgnug\textsuperscript{1},
Lars Eric Kroll\textsuperscript{1}%,
%Name4 Surname\textsuperscript{2},
%Name5 Surname\textsuperscript{2\ddag},
%Name6 Surname\textsuperscript{2\ddag},
%Name7 Surname\textsuperscript{1,2,3*},
%with the Lorem Ipsum Consortium\textsuperscript{\textpilcrow}
\\
\bigskip
\textbf{1} Central Research Institute of Ambulatory Health Care in Germany (Zi),
Salzufer 8, D-10587 Berlin, Germany
\\
%\textbf{2} Affiliation Dept/Program/Center, Institution Name, City, State, Country
%\\
%\textbf{3} Affiliation Dept/Program/Center, Institution Name, City, State, Country
%\\
\bigskip

@@

We used a directed acyclic graph (DAG)
\cite{schipf_directed_2011, textor_robust_2017} as a tool to analyze
the causal relationships between several exposures and SARS-CoV-2
spread. To get an overview on published associations, a scoping review
was conducted from 20th to 22nd of May 2020 within Pubmed and Google
scholar. Restrictions were applied to English and German language and
the publication date in the last one year. The following search terms
were applied to abstracts and title in Pubmed (``COVID-19'' OR
``COVID19'' OR ``Corona'' OR ``Coronavirus'' OR ``SARS-CoV-2'') and
connected separately in each case with the exposure variables
(``mobility'', ``public awareness'', ``awareness'', ``google
trends'',``ambient temperature'', ``temperature''). For ``mobility'', we
analyzed \(n=8\) studies, \(N=103\) were scanned in Pubmed, together
with the first ten pages (100 results) in Google scholar
(``awareness''/``public awareness''/``google trends'' \(n=9\),
\(N=215\); ``temperature''/``ambient temperature'' \(n=16\), \(N=235\)).
We integrated these findings where possible into the construction of our
DAG, which can be seen in Fig~\ref{fig1}.

@@

\begin{figure}[!h]
\caption{{\bf DAG of determinants of reported COVID-19 cases on the district level.}
Unobserved variables are light gray, variables marked with an asterisk (*) are confounded by weekday/holiday.}
\label{fig1}
\end{figure}

@@

We collected and aggregated data on reported COVID-19 cases, regional
socio-demographic factors, weather, and general mobility on district and
state level in Germany for the period of 15 February 2020 to 8 July
2020. Our observation period for the outcome consisted of all dates from
20 February 2020 to 8 July 2020 (\(T=140\)), since we used a lag of 5
days for all confounders. We did not exclude any states or districts
(\(K=401\)). We analyzed the daily reported number of new cases as
outcome (\(K\cdot T=56{\ }140\) observations). The set of possible
predictors was derived from our causal DAG (see Table~\ref{tab:model-vars} and Fig~\ref{fig1}). Due to
modelling and data limitations, some of the predictors were unobserved
or were modelled as a construct consisting of several variables. For our causal graph analysis, we computed adjustment sets separately for all observed exposures within the DAG (if the respective exposure was identifiable within the DAG causal analysis framework).

@@

We analyzed the DAG from Fig~\ref{fig1} with the R Software
\cite{rsoftware} and the R packages \texttt{dagitty} (formal
representation of the graph and minimal adjustment sets
\cite{textor_robust_2017}) and \texttt{pcalg} (for finding an optimal
adjustment set \cite{pcalg}). For the defined exposures and the outcome
``Reported new cases of COVID-19'', we computed the minimal and optimal
adjustment sets. Since it was possible that these sets contained unobserved variables that needed to be left out of the regression model, we chose the valid set with the lowest AIC (see next section) to estimate the final total causal effect from exposure to outcome.

@@

In the observational period, the number of daily reported COVID-19 cases increased till the end of March/beginning of April and continually decreased afterwards till the beginning of June 2020 with a slight increase and decrease afterwards (Fig~\ref{fig2}A). On the other hand, the (log-)ratio of reported cases over active cases decreased steeply till the mid of April and increased steadily afterwards with a slight decrease close to the end of the observation period (Fig~\ref{fig2}B). Both figures examplify a considerable variation among the districts (light blue points are individual district's data).

\begin{figure}[!h]
\caption{{\bf Temporal and district level variation of outcome (log-scale).}}
\label{fig2}
\end{figure}

In Germany, we observed a rebound in mobility after the initial political measures, reductions in incident cases were associated with a diminishing public interest in COVID-19, and temperatures were overall increasing (cf.~Fig~\ref{fig3}); with correlations between temporal progression and mobility in retail and recreation \(r_{A,B}=\) 0.02, awareness (``Searches corona'') \(r_{A,C}=\) -0.3, and temperature \(r_{A,D}=\) 0.8.

\begin{figure}[!h]
\caption{{\bf Temporal variation of outcome and main
determinants.}}
\label{fig3}
\end{figure}

@@

We list the results of our causal analysis for the effects of different exposure variables in Table~\ref{tab:causaleffects}. The estimates are multiplicative rates of increase/decrease for a one unit increase of the respective variable: Values above 1 lead to an increase, below 1 to a decrease of the infection rate. To put these estimates into perspective, Fig~\ref{fig4} shows the relative causal effect of the different exposure variables on the number of reported COVID-19 cases on a range of sensible values of the exposure variables (95 percent quantiles of data points).

@@

\begin{figure}[!h]
\caption{{\bf Relative causal effects of exposures.}}
\label{fig4}
\end{figure}

Within our framework, we saw very different effects for individual mobility variables. For mobility in retail/recreation, an increase of 1 percent point mobility compared to the reference period (03 January to 06 February 2020) leads to an increase of the daily reported case number by about 0.11 percent. Similarly, mobility on workplaces showed an effect of 0.33 increase in case numbers for every 1 percent point increase in mobility, while mobility on transit stations showed an effect of 0.26 increase in case numbers for every 1 percent point increase. Contrarily, the remaining three mobility variables showed negative effects on the number of reported COVID-19 cases. An increase of 1 percent point mobility for the areas of grocery/pharmacy leads to a decrease in the reported case number by approximately 0.23 percent, while increased mobility of 1 percent point within parks leads to a decrease in the reported case number by approximately 0.03 percent, and finally an increase of 1 percent point in residential mobility leads to a decrease by approximately 0.97 percent. Fig~\ref{fig4} shows the effects of mobility on a range of possible values. Thus, we expect an increase of daily cases by approximately 7.8 percent if mobility in workplaces reaches baseline levels of 0 percent difference to the reference period. On the other hand, an increase of mobility for residential areas by 10 percent points compared to the reference period leads to a reduction of the infection rate by approximately 1.8 percent.

``Awareness'' had two opposite effects on the outcome in our DAG. Awareness measured by Google searches for \emph{corona} had a positive effect on the number of reported cases. An one percent point increase of the state's Google searches (relative to other states and the observation period) leads to an increase of approximately 0.89 percent. For example, if a district shows 10 percent points more relative searches for \emph{corona} than another one, we expect approximately 9.3 percent more infections for this district after 5 days. \emph{COVID-19 burden} (reported number of cases on day of exposure) affected the outcome negatively, where every additional daily case in the district leads to a 0.2 percent decrease in newly reported case numbers. The corresponding plot in Fig~\ref{fig4} visualizes this relationship: For a local outbreak with 20 daily cases as COVID-19 burden, we estimate as total causal effect a subsequent reduction of infection rate by 3.9 percent.

Within our model, we observed effects of temperature and all other weather variables. Every increase of 1 degree Celsius in temperature leads to a reduction of the daily reported case numbers by approximately 0.95 percent. On the other hand, we found an increasing effect of rainfall: One millimeter (=1 liter per square meter) more rainfall leads to an increase of reported case numbers by approximately 1.21 percent. We observe effects for humidity and wind as well (higher humidity and stronger wind leading to more cases). In perspective (Fig~\ref{fig4}), with temperature we expect an increase by approximately 21 percent at a daily average temperature of \(0^{\circ}C\) compared to a day with \(20^{\circ}C\). For rainfall, we expect on a rainy day with 10 mm rainfall a corresponding increase of the infection rate by approximately 12.8 percent compared to a day with no precipitation.

@@

The code and resources for our analysis are available on Github, we invite other researchers to replicate our analysis with different assumptions using the files provided in the repository of the article (\url{https://github.com/zidatalab/causalcovid19}).

@@

%\bibliography{mybibfile}

\begin{thebibliography}{10}

\bibitem{who2020report}
{WHO Team}. Report of the {WHO-China} Joint Mission on Coronavirus Disease 2019
  ({COVID-19}); 2020, accessed 2020-06-25.
\newblock Available from:
  \url{https://www.who.int/publications-detail/report-of-the-who-china-joint-mission-on-coronavirus-disease-2019-(covid-19)}.

\bibitem{Chinazzi395}
Chinazzi M, Davis JT, Ajelli M, Gioannini C, Litvinova M, Merler S, et~al.
\newblock The effect of travel restrictions on the spread of the 2019 novel
  coronavirus ({COVID-19}) outbreak.
\newblock Science. 2020;368(6489):395--400.
\newblock doi:{10.1126/science.aba9757}.

\bibitem{guan2020clinical}
Guan W, Ni Z, Hu Y, Liang W, Ou C, He J, et~al.
\newblock Clinical Characteristics of Coronavirus Disease 2019 in {China}.
\newblock New England Journal of Medicine. 2020;doi:{10.1056/NEJMoa2002032}.

\bibitem{higgins_correlations_2020}
Higgins TS, Wu AW, Sharma D, Illing EA, Rubel K, Ting JY.
\newblock Correlations of {Online} {Search} {Engine} {Trends} {With}
  {Coronavirus} {Disease} ({COVID}-19) {Incidence}: {Infodemiology} {Study}.
\newblock JMIR Public Health and Surveillance. 2020;6(2):e19702.
\newblock doi:{10.2196/19702}.

\bibitem{li_retrospective_2020}
Li C, Chen LJ, Chen X, Zhang M, Pang CP, Chen H.
\newblock Retrospective analysis of the possibility of predicting the
  {COVID}-19 outbreak from {Internet} searches and social media data, {China},
  2020.
\newblock Euro Surveillance. 2020;25(10).
\newblock doi:{10.2807/1560-7917.ES.2020.25.10.2000199}.

\bibitem{yuan_trends_2020}
Yuan X, Xu J, Hussain S, Wang H, Gao N, Zhang L.
\newblock Trends and {Prediction} in {Daily} {New} {Cases} and {Deaths} of
  {COVID}-19 in the {United} {States}: {An} {Internet} {Search}-{Interest}
  {Based} {Model}.
\newblock Exploratory research and hypothesis in medicine. 2020;5(2):1--6.
\newblock doi:{10.14218/ERHM.2020.00023}.

\bibitem{bannister-tyrrell_preliminary_2020}
Bannister-Tyrrell M, Meyer A, Faverjon C, Cameron A.
\newblock Preliminary evidence that higher temperatures are associated with
  lower incidence of {COVID}-19, for cases reported globally up to 29th
  {February} 2020.
\newblock medRxiv. 2020;doi:{10.1101/2020.03.18.20036731}.

\bibitem{demongeot_temperature_2020}
Demongeot J, Flet-Berliac Y, Seligmann H.
\newblock Temperature {Decreases} {Spread} {Parameters} of the {New} {Covid}-19
  {Case} {Dynamics}.
\newblock Biology. 2020;9(5).
\newblock doi:{10.3390/biology9050094}.

\bibitem{liu_impact_2020}
Liu J, Zhou J, Yao J, Zhang X, Li L, Xu X, et~al.
\newblock Impact of meteorological factors on the {COVID}-19 transmission: {A}
  multi-city study in {China}.
\newblock Science of the Total Environment. 2020;726:138513.
\newblock doi:{10.1016/j.scitotenv.2020.138513}.

\bibitem{Greenland1999}
Greenland S, Robins JM, Pearl J.
\newblock Confounding and {Collapsibility} in {Causal} {Inference}.
\newblock Statistical Science. 1999;14(1):29--46.

\bibitem{schipf_directed_2011}
Schipf S, Knüppel S, Hardt J, Stang A.
\newblock Directed {Acyclic} {Graphs} ({DAGs}) – {Die} {Anwendung} kausaler
  {Graphen} in der {Epidemiologie}.
\newblock Gesundheitswesen. 2011;73(12):888--892.
\newblock doi:{10.1055/s-0031-1291192}.

\bibitem{textor_robust_2017}
Textor J, van~der Zander B, Gilthorpe MS, Liśkiewicz M, Ellison GT.
\newblock Robust causal inference using directed acyclic graphs: the {R}
  package ‘dagitty’.
\newblock International Journal of Epidemiology. 2017;45(6):1887--1894.
\newblock doi:{10.1093/ije/dyw341}.

\bibitem{jhucovid19db2020}
{Center for Systems Science and Engineering (CSSE)}. {COVID-19 Data Repository
  by the Center for Systems Science and Engineering (CSSE) at Johns Hopkins
  University}; 2020.
\newblock Available from: \url{https://github.com/CSSEGISandData/COVID-19}.

\bibitem{Pearl_2014}
Pearl J, Bareinboim E.
\newblock External Validity: From Do-Calculus to Transportability Across
  Populations.
\newblock Statistical Science. 2014;29(4):579–595.
\newblock doi:{10.1214/14-sts486}.

\bibitem{chang_modeling_2020}
Chang MC, Kahn R, Li YA, Lee CS, Buckee CO, Chang HH.
\newblock Variation in human mobility and its impact on the risk of future
  {COVID}-19 outbreaks in {Taiwan}.
\newblock BMC Public Health. 2021;21(1):226.
\newblock doi:{10.1186/s12889-021-10260-7}.

\bibitem{fowler_effect_2020}
Fowler JH, Hill SJ, Obradovich N, Levin R.
\newblock {The Effect of Stay-at-Home Orders on COVID-19 Cases and Fatalities
  in the United States}.
\newblock medRxiv. 2020;doi:{10.1101/2020.04.13.20063628}.

\bibitem{kraemer_effect_2020}
Kraemer MUG, Yang CH, Gutierrez B, Wu CH, Klein B, Pigott DM, et~al.
\newblock The effect of human mobility and control measures on the {COVID}-19
  epidemic in {China}.
\newblock Science (New York, NY). 2020;368(6490):493--497.
\newblock doi:{10.1126/science.abb4218}.

\bibitem{lasry_timing_2020}
Lasry A, Kidder D, Hast M, Poovey J, Sunshine G, Winglee K, et~al.
\newblock Timing of Community Mitigation and Changes in Reported {COVID-19} and
  Community Mobility - Four {U.S.} Metropolitan Areas, {February 26-April 1,
  2020}.
\newblock MMWR Morbidity and mortality weekly report. 2020;69(15):451—457.
\newblock doi:{10.15585/mmwr.mm6915e2}.

\bibitem{linka_outbreak_2020}
Linka K, Peirlinck M, Sahli~Costabal F, Kuhl E.
\newblock Outbreak dynamics of {COVID}-19 in {Europe} and the effect of travel
  restrictions.
\newblock Computer Methods in Biomechanics and Biomedical Engineering. 2020; p.
  1--8.
\newblock doi:{10.1080/10255842.2020.1759560}.

\bibitem{mazzoli_effects_2020}
Mazzoli M, Mateo D, Hernando A, Meloni S, Ramasco JJ.
\newblock Effects of mobility and multi-seeding on the propagation of the
  {COVID}-19 in {Spain}.
\newblock medRxiv. 2020;doi:{10.1101/2020.05.09.20096339}.

\bibitem{xiong_data-driven_2020}
{Xiong} C, {Hu} S, {Yang} M, {Younes} HN, {Luo} W, {Ghader} S, et~al.
\newblock {Data-Driven Modeling Reveals the Impact of Stay-at-Home Orders on
  Human Mobility during the COVID-19 Pandemic in the U.S}.
\newblock arXiv e-prints. 2020; p. arXiv:2005.00667.

\bibitem{juni_impact_2020}
Jüni P, Rothenbühler M, Bobos P, Thorpe KE, da~Costa BR, Fisman DN, et~al.
\newblock Impact of climate and public health interventions on the {COVID}-19
  pandemic: {A} prospective cohort study.
\newblock Canadian Medical Association Journal. 2020;doi:{10.1503/cmaj.200920}.

\bibitem{lai_effect_2020}
Lai S, Ruktanonchai NW, Zhou L, Prosper O, Luo W, Floyd JR, et~al.
\newblock Effect of non-pharmaceutical interventions to contain {COVID-19} in
  {China}.
\newblock Nature. 2020;doi:{10.1038/s41586-020-2293-x}.

\bibitem{google_trends}
{Google LLC}. {Google Trends, search term {``}corona{''}}; 2020, accessed
  2020-06-25.
\newblock Available from: \url{https://www.google.com/trends}.

\bibitem{ayyoubzadeh_predicting_2020}
Ayyoubzadeh SM, Ayyoubzadeh SM, Zahedi H, Ahmadi M, R~Niakan~Kalhori S.
\newblock Predicting {COVID}-19 {Incidence} {Through} {Analysis} of {Google}
  {Trends} {Data} in {Iran}: {Data} {Mining} and {Deep} {Learning} {Pilot}
  {Study}.
\newblock JMIR Public Health and Surveillance. 2020;6(2):e18828.
\newblock doi:{10.2196/18828}.

\bibitem{effenberger_association_2020}
Effenberger M, Kronbichler A, Shin JI, Mayer G, Tilg H, Perco P.
\newblock Association of the {COVID}-19 pandemic with {Internet} {Search}
  {Volumes}: {A} {Google} {Trends}({TM}) {Analysis}.
\newblock International Journal of Infectious Diseases. 2020;95:192--197.
\newblock doi:{10.1016/j.ijid.2020.04.033}.

\bibitem{lin_google_2020}
Lin YH, Liu CH, Chiu YC.
\newblock Google searches for the keywords of ``wash hands'' predict the speed
  of national spread of {COVID}-19 outbreak among 21 countries.
\newblock Brain, Behavior, and Immunity. 2020;doi:{10.1016/j.bbi.2020.04.020}.

\bibitem{mavragani_tracking_2020}
Mavragani A.
\newblock Tracking {COVID}-19 in {Europe}: {Infodemiology} {Approach}.
\newblock JMIR Public Health and Surveillance. 2020;6(2):e18941.
\newblock doi:{10.2196/18941}.

\bibitem{walker_use_2020}
Walker A, Hopkins C, Surda P.
\newblock Use of {Google Trends} to investigate loss-of-smell-related searches
  during the {COVID-19} outbreak.
\newblock International Forum of Allergy \& Rhinology. 2020;10(7):839--847.
\newblock doi:{10.1002/alr.22580}.

\bibitem{zhou_effects_2020}
Zhou WK, Wang AL, Xia F, Xiao YN, Tang SY.
\newblock Effects of media reporting on mitigating spread of {COVID}-19 in the
  early phase of the outbreak.
\newblock Mathematical Biosciences and Engineering. 2020;17(3):2693--2707.
\newblock doi:{10.3934/mbe.2020147}.

\bibitem{qi_covid-19_2020}
Qi H, Xiao S, Shi R, Ward MP, Chen Y, Tu W, et~al.
\newblock {COVID}-19 transmission in {Mainland} {China} is associated with
  temperature and humidity: {A} time-series analysis.
\newblock Science of the Total Environment. 2020;728:138778.
\newblock doi:{10.1016/j.scitotenv.2020.138778}.

\bibitem{shi_impact_2020}
Shi P, Dong Y, Yan H, Zhao C, Li X, Liu W, et~al.
\newblock Impact of temperature on the dynamics of the {COVID}-19 outbreak in
  {China}.
\newblock Science of the Total Environment. 2020;728:138890.
\newblock doi:{10.1016/j.scitotenv.2020.138890}.

\bibitem{sobral_association_2020}
Sobral MFF, Duarte GB, da~Penha~Sobral AIG, Marinho MLM, de~Souza~Melo A.
\newblock Association between climate variables and global transmission {of}
  {SARS}-{CoV}-2.
\newblock Science of the Total Environment. 2020;729:138997.
\newblock doi:{10.1016/j.scitotenv.2020.138997}.

\bibitem{tosepu_correlation_2020}
Tosepu R, Gunawan J, Effendy DS, Ahmad LOAI, Lestari H, Bahar H, et~al.
\newblock Correlation between weather and {Covid}-19 pandemic in {Jakarta},
  {Indonesia}.
\newblock Science of the Total Environment. 2020;725:138436.
\newblock doi:{10.1016/j.scitotenv.2020.138436}.

\bibitem{Wang2020temperature}
Wang M, Jiang A, Gong L, Luo L, Guo W, Li C, et~al.
\newblock Temperature significant change {COVID-19} Transmission in 429 cities.
\newblock medRxiv. 2020;doi:{10.1101/2020.02.22.20025791}.

\bibitem{wu_effects_2020}
Wu Y, Jing W, Liu J, Ma Q, Yuan J, Wang Y, et~al.
\newblock Effects of temperature and humidity on the daily new cases and new
  deaths of {COVID-19} in 166 countries.
\newblock Science of the Total Environment. 2020;729:139051.
\newblock doi:{https://doi.org/10.1016/j.scitotenv.2020.139051}.

\bibitem{auler_evidence_2020}
Auler AC, Cássaro FAM, da~Silva VO, Pires LF.
\newblock Evidence that high temperatures and intermediate relative humidity
  might favor the spread of {COVID}-19 in tropical climate: {A} case study for
  the most affected {Brazilian} cities.
\newblock Science of the Total Environment. 2020;729:139090.
\newblock doi:{10.1016/j.scitotenv.2020.139090}.

\bibitem{xie_association_2020}
Xie J, Zhu Y.
\newblock Association between ambient temperature and {COVID}-19 infection in
  122 cities from {China}.
\newblock Science of the Total Environment. 2020;724:138201.
\newblock doi:{10.1016/j.scitotenv.2020.138201}.

\bibitem{briz-redon_spatio-temporal_2020}
Briz-Red{\'o}n {\'A}, Serrano-Aroca {\'A}.
\newblock A spatio-temporal analysis for exploring the effect of temperature on
  {COVID}-19 early evolution in {Spain}.
\newblock Science of the Total Environment. 2020;728:138811.
\newblock doi:{10.1016/j.scitotenv.2020.138811}.

\bibitem{iqbal_nexus_2020}
Iqbal N, Fareed Z, Shahzad F, He X, Shahzad U, Lina M.
\newblock The nexus between {COVID}-19, temperature and exchange rate in
  {Wuhan} city: {New} findings from partial and multiple wavelet coherence.
\newblock Science of the Total Environment. 2020;729:138916.
\newblock doi:{10.1016/j.scitotenv.2020.138916}.

\bibitem{jahangiri_sensitivity_2020}
Jahangiri M, Jahangiri M, Najafgholipour M.
\newblock The sensitivity and specificity analyses of ambient temperature and
  population size on the transmission rate of the novel coronavirus
  ({COVID}-19) in different provinces of {Iran}.
\newblock Science of the Total Environment. 2020;728:138872.
\newblock doi:{10.1016/j.scitotenv.2020.138872}.

\bibitem{yao_no_2020}
Yao Y, Pan J, Liu Z, Meng X, Wang W, Kan H, et~al.
\newblock No association of {COVID}-19 transmission with temperature or {UV}
  radiation in {Chinese} cities.
\newblock The European Respiratory Journal. 2020;55(5).
\newblock doi:{10.1183/13993003.00517-2020}.

\bibitem{de2020risk}
de~Lusignan S, Dorward J, Correa A, Jones N, Akinyemi O, Amirthalingam G,
  et~al.
\newblock Risk factors for {SARS}-{CoV}-2 among patients in the {Oxford}
  {Royal} {College} of {General} {Practitioners} {Research} and {Surveillance}
  {Centre} primary care network: a cross-sectional study.
\newblock The Lancet Infectious Diseases;doi:{10.1016/S1473-3099(20)30371-6}.

\bibitem{Dragano2020.06.17.20133918}
Wahrendorf M, Rupprecht CJ, Dortmann O, Scheider M, Dragano N.
\newblock Erhöhtes {Risiko} eines {COVID}-19-bedingten
  {Krankenhausaufenthaltes} für {Arbeitslose}: {Eine} {Analyse} von
  {Krankenkassendaten} von 1,28 {Mio}. {Versicherten} in {Deutschland}.
\newblock Bundesgesundheitsblatt - Gesundheitsforschung - Gesundheitsschutz.
  2021;64(3):314--321.
\newblock doi:{10.1007/s00103-021-03280-6}.

\bibitem{dohle_wingen_schreiber_2020}
Dohle S, Wingen T, Schreiber M.
\newblock Acceptance and Adoption of Protective Measures During the {COVID}-19
  Pandemic: The Role of Trust in Politics and Trust in Science.
\newblock Social Psychological Bulletin. 2020;15(4):1--23.
\newblock doi:{10.32872/spb.4315}.

\bibitem{engle_staying_2020}
Engle S, Stromme J, Zhou A.
\newblock Staying at home: mobility effects of {COVID-19}.
\newblock Available at SSRN. 2020;doi:{10.2139/ssrn.3565703}.

\bibitem{cowling2020impact}
Cowling BJ, Ali ST, Ng TWY, Tsang TK, Li JCM, Fong MW, et~al.
\newblock Impact assessment of non-pharmaceutical interventions against
  coronavirus disease 2019 and influenza in {Hong} {Kong}: an observational
  study.
\newblock The Lancet Public Health. 2020;5(5):e279--e288.
\newblock doi:{10.1016/S2468-2667(20)30090-6}.

\bibitem{casenumbers_rki}
{Robert Koch-Institut (RKI)}. {Fallzahlen in Deutschland (COVID-19)}; 2020,
  accessed 2020-07-12.
\newblock Available from:
  \url{https://www.rki.de/DE/Content/InfAZ/N/Neuartiges_Coronavirus/Fallzahlen.html}.

\bibitem{google_mobility}
{Google LLC}. Google {COVID-19} Community Mobility Reports; 2020, accessed
  2020-06-25.
\newblock Available from: \url{https://www.google.com/covid19/mobility/}.

\bibitem{dwd_weather}
{Deutscher Wetterdienst (DWD) Climate Data Center (CDC)}. Recent daily station
  observations (temperature, pressure, precipitation,sunshine duration, etc.)
  for {Germany}, quality control not completed yet, version recent; 2020,
  accessed 2020-07-12.
\newblock Available from:
  \url{https://opendata.dwd.de/climate_environment/CDC/observations_germany/climate/daily/kl/recent/}.

\bibitem{inkar}
{Bundesinstitut für Bau-, Stadt- und Raumforschung (BBSR)}. {INKAR –
  Indikatoren und Karten zur Raum- und Stadtentwicklung}; 2020, accessed
  2020-06-25.
\newblock Available from: \url{https://www.inkar.de/}.
